\subsection{Clang}{Delio Vicini}

Clang is an open-source compiler for C, C++, Objective-C, Objective-C++, and CUDA. The Clang compiler project was started by Apple in 2005 in order to replace the GNU Compiler Collection (GCC). The original motivation for the Clang project was to have a fast, modular compiler that is easier to extend than GCC and cooperates well with integrated development environments (IDEs).\cite{clang-motivation} A tight integration of the compiler with the IDE may be used to allow things such as code auto-completion using the data structures used by the compiler itself. Technically, the Clang compiler builds on the LLVM\footnote{LLVM is not an abbreviation and is the full name of the project. Historically, it used to be the abbrevation for "Low-Level Virtual Machine."} compiler infrastructure and is a sub-project thereof. Both Clang and LLVM were originally developed by Chris Lattner.\cite{lattner} \\

The Clang compiler is licensed under the University of Illinois/NCSA License.\cite{clang-policy} This is a fairly permissive license, comparable to the BSD family of licenses. It is more permissive than a GPL-style copyleft license, as it does not require publishing changes made to the original source code. The redistributed software merely needs to contain the original licensing text, and must not be endorsed using the names or affiliations of the original authors. Furthermore, no warranty for correct or safe execution is given.\cite{illinois-license} \\

Being part of the LLVM project, Clang is also governed by the not-for-profit LLVM Foundation. The foundation has the goal of providing educational training and resources relating to the LLVM project and technology.\cite{llvm-foundation} The board of the foundation however typically does not decide on code changes and patches. All low-level technical aspects are directly handled by the community.\cite{clang-policy} \\

Contributions from the community are accepted traditionally via a developer mailing list, similar to the procedure that was described in the previous section on Linux. Everyone can submit patches to this mailing list and can comment on other contributions. However, only people with write access to the repository can approve patches and merge them into the main repository. Patches are typically reviewed before being committed to the repository. Post-commit review is an option for trusted maintainers, allowing them more rapid development. In order to obtain write access for the repository, one has to have a track record of high quality contributions, and send a request to Chris Lattner, the original creator of Clang. \\

To guarantee that all code is reviewed, even in the case of post-commit review, there is a code-owner associated to each piece of code in the repository. Code-owners are responsible for the quality control of their respective area of the code base. In addition to this, quality is enforced via detailed coding standards, unit tests, and automated nightly builds. \\

As Clang was originally started by Apple, it is no surprise that Apple is still heavily involved in its development. Chris Lattner is still an active member of the LLVM Foundation board and still works for Apple.\cite{lattner} Furthermore, the five people with the most commits to the Clang project are all Apple employees as well.\cite{clang-commits} The situation is a bit more diverse on the LLVM Foundation board, where employees from ARM, Google, Qualcomm, and academia can be found.\cite{llvm-board} The LLVM Foundation is sponsored by Apple, Google, Facebook, Intel, and a few smaller companies.\cite{llvm-sponsors} \\

Despite the strong industry involvement, it seems that the development of Clang is mostly driven by the community, and less so by the LLVM board and the involed companies. According to the official development policy, conflicts should be resolved by discussion via the appropriate mailing list.\cite{clang-policy} Given the permissive license of Clang, it is very likely that some companies maintain an internal version of Clang, which is more tailored towards their respective needs. The most prominent example is Apple, where many new compiler features are first developed internally before they are potentially merged into the open-source Clang repository. For example, the Clang version distributed with Apple's XCode IDE is not the open-source version. This lessens the chance of big conflicts in the development of Clang.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "foss-governance"
%%% End:

