\subsection{PBRT}{Felice Serena}

% Intro, History
PBRT stands for ``Physically Based Ray Tracer''. 
It is a render engine described in the book ``Physically Based Rendering: 
From Theory to Implementation'' by Matt Pharr, Wenzel Jakob, and Greg Humphreys. 
They open-sourced their implementation to provide a complete application for students and researchers to study. 
PBRT is available under the BSD license. 
\\

% Community
The authors Matt Pharr, Wenzel Jakob, and Greg Humphreys stand at the top of the community hierarchy and make all decisions amongst them. 
They made it very clear that they wanted to keep the software close to the book. 
The reason is that the entire source code is written down and explained in the book. 
If one reads the entire book, he/she also has read the whole source code of PBRT.
Because of this tight coupling, the authors rarely accept pull-requests for new features. 
Though they encourage people to create more unit tests and scene-converters, improve portability or push bug fixes. \cite{pbrt-pull-requests} \\

% companies
Since PBRT is so restrictive to new features, no companies are involved in its development. \\

% conflicts
Conflicts are evident. 
The authors intend to sell their book and so keep the source code close to the printed original. 
A big part of the community wants to build on top of the existing engine and add more features. 
This conflict forked a new open-source project: LuxRender \cite{luxrender-home}.
LuxRender aims to be an unbiased, physically based render engine with steady development.
\\

% code
The code base is mainly maintained by the three authors.
Though, the repository lists 16 other contributors.
The primary language is C++; git is used as version control system.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "foss-governance"
%%% End:
