\section{Common Aspects}
Despite the diversity in the projects, we have been able to determine several common aspects. The majority of the projects were started by a single person, who had a strong personal motivation to get started. The community then branched out from there, attracting people with similar goals to contribute. The usual structure found in the projects also reflects this, where a very select number of people sit at the top of the hierarchy and make most of the important decisions and delegate smaller ones to maintainers. \\

However, the actual course of the project is usually determined by the drive of the community, as the only way to enforce advancement is to contribute yourself. The reasons that drive people to contribute can be quite varied, with the most common being the need to use the software for their own purpose, or wanting to further their knowledge, career, and community involvement. The quality assurance is also handled through the community hierarchy, where people higher up take on positions that are much more strongly tending towards review and discussion, rather than the writing of new code. \\

Corporate involvement usually seems to stem from the company's reliance on the software for their own products, particularly when they are a hardware vendor that needs tailored support. However, it is also the case that corporations will open-source their own software in order to outsource the necessary work to continue its development. Allowing community contributions can improve a product massively without the need to pay more employees. \\

Unlike closed-source software, open-source projects are able to sustain themselves over very long times to the point where it is practically impossible for one to die out. Anyone with a copy of the source code could revive the project at any time, or split it off into a new project with different goals and aims. Closed-source projects do not have this advantage, as their license restrictions would prevent other people from continuing work, not to mention that usually the code itself is unavailable completely and it would be impractical to reverse-engineer it.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "foss-governance"
%%% End:
